---
title: "recreation of: An Empirical Study of Model-Agnostic Techniques for Defect Prediction Models"
output: html_notebook
---

```{r}
library(Rnalytica)
library(tidyverse)
library(tidymodels)
library(lime)
library(DALEX)
```

# data

```{r}
defects <- loadDefectDataset('activemq-5.0.0', 'jira')
data <- defects$data %>% drop_na()
```

Split. 

```{r}
set.seed(123)

data_split <- initial_split(data, strata = RealBug)
data_train <- training(data_split)
data_test <- testing(data_split)
```


```{r}
first_sample <- select(data_test, -RealBug) %>% slice(1)
first_sample
```

```{r}
second_sample <- select(data_test, -RealBug) %>% slice(2)
second_sample
```

Bootstrap resamples.

```{r}
set.seed(123)

folds <- bootstraps(data_train, strata = RealBug, apparent = FALSE, times = 100)
```


# formula construction

Independent variables should be the same for each dataset.

```{r}
indep <- defects$indep
indep
```

Paper mentions using AutoSpearman to pick non-correlated features. 

```{r}
features <- AutoSpearman(dataset = data, metrics = indep)

features
```

```{r}
formula <- RealBug ~ CountDeclMethodPrivate + CountDeclMethodDefault + AvgEssential + CountDeclClassVariable + CountDeclClassMethod + AvgLineComment + AvgCyclomaticModified + CountDeclClass +
  CountDeclMethodProtected + CountDeclInstanceVariable + RatioCommentToCode + AvgLineBlank + PercentLackOfCohesion + PercentLackOfCohesion + MaxInheritanceTree + CountClassDerived + CountClassCoupled + 
  CountClassBase + CountInput_Mean + CountInput_Min + CountOutput_Mean + CountOutput_Min + CountPath_Min + MaxNesting_Min + OWN_COMMIT + MINOR_COMMIT + MINOR_LINE + MAJOR_LINE
```

```{r}
rec <- recipe(formula, data_train)
```


# random forest model

## workflow

```{r}
rf_mod <-
  rand_forest(trees = 64) %>%
  set_engine("ranger") %>%
  set_mode("classification")
```

```{r}
rf_wf <- 
  workflow() %>% 
  add_model(rf_mod) %>%
  add_recipe(rec)
```

## fit resample

```{r}
set.seed(123)
rf_fit_rs <- 
  rf_wf %>%
  fit_resamples(folds)
```

 
```{r}
collect_metrics(rf_fit_rs)
```

The resample suggests that the model is a good classifier. 

## fit model

Create the model normally. The type flag gives the probabilities for each class: "TRUE" or "FALSE".

```{r}
rf_fit <- fit(rf_wf, data=data_train)
rf_pred <- predict(rf_fit, data_test, type = 'prob')
```

## prediction metrics

```{r}
rf_pred <- bind_cols(truth=data_test$RealBug, rf_pred)
rf_pred
```

```{r}
roc_curve(
  rf_pred,
  truth,
  .pred_TRUE,
  event_level = "second"
) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity)) +
  geom_path() +
  geom_abline(lty = 3) +
  coord_equal() +
  theme_bw()
```

ROC curve shows that our model does an alright job in its predictions. The curve is similar to the roc_auc value we found while resampling. 


## explainations

```{r}
model <- extract_fit_parsnip(rf_fit)
```

### lime explaination

```{r}
explainer <- lime(select(data_train, -RealBug), model, bin_continuous = TRUE, quantile_bins = FALSE)

first_explaination <- lime::explain(first_sample, explainer=explainer, n_features = 5, n_labels=2)

first_explaination
```


```{r}
plot_features(first_explaination)
```

```{r}
second_explaination <- lime::explain(second_sample, explainer=explainer, n_features = 5, n_labels=2)

second_explaination
```

```{r}
plot_features(second_explaination)
```

Although only slightly, the explanations between two samples ARE different. Compare CountClassCoupled between the LIME explainners. 

### breakdown explanation

Breakdown of the first sample:

```{r}
DALEX::explain(model, data = select(data_train, -RealBug), y = data_train$RealBug) %>%
  predict_parts(new_observation = first_sample, type = "break_down") %>% 
  plot()
```

Breakdown of the second sample:

```{r}
DALEX::explain(model, data = select(data_train, -RealBug), y = data_train$RealBug) %>%
  predict_parts(new_observation = second_sample, type = "break_down") %>% 
  plot()
```

# logistic model

## workflow

```{r}
log_mod <-
  logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")
```

```{r}
rf_wf <- 
  workflow() %>% 
  add_model(log_mod) %>%
  add_recipe(rec)
```

## fit resample

```{r}
set.seed(123)
rf_fit_rs <- 
  rf_wf %>%
  fit_resamples(folds)
```

 
```{r}
collect_metrics(rf_fit_rs)
```


## fit model


```{r}
rf_fit <- fit(rf_wf, data=data_train)
rf_pred <- predict(rf_fit, data_test, type = 'prob')
```

## prediction metrics

```{r}
rf_pred <- bind_cols(truth=data_test$RealBug, rf_pred)
rf_pred
```

```{r}
roc_curve(
  rf_pred,
  truth,
  .pred_TRUE,
  event_level = "second"
) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity)) +
  geom_path() +
  geom_abline(lty = 3) +
  coord_equal() +
  theme_bw()
```

## explainations

```{r}
model <- extract_fit_parsnip(rf_fit)
```

### lime explaination

```{r}
explainer <- lime(select(data_train, -RealBug), model, bin_continuous = TRUE, quantile_bins = FALSE)

first_explaination <- lime::explain(first_sample, explainer=explainer, n_features = 5, n_labels=2)

first_explaination
```


```{r}
plot_features(first_explaination)
```

```{r}
second_explaination <- lime::explain(second_sample, explainer=explainer, n_features = 5, n_labels=2)

second_explaination
```

```{r}
plot_features(second_explaination)
```

Although only slightly, the explanations between two samples ARE different. Compare CountClassCoupled between the LIME explainners. 

### breakdown explanation

Breakdown of the first sample:

```{r}
DALEX::explain(model, data = select(data_train, -RealBug), y = data_train$RealBug) %>%
  predict_parts(new_observation = first_sample, type = "break_down") %>% 
  plot()
```

Breakdown of the second sample:

```{r}
DALEX::explain(model, data = select(data_train, -RealBug), y = data_train$RealBug) %>%
  predict_parts(new_observation = second_sample, type = "break_down") %>% 
  plot()
```


# reference
 
https://rsample.tidymodels.org/reference/bootstraps.html
https://www.tidymodels.org/start/resampling/
https://yardstick.tidymodels.org/reference/roc_curve.html
https://medium.com/responsibleml/basic-xai-with-dalex-part-6-lime-method-f6aab0af058a
https://yardstick.tidymodels.org/reference/roc_curve.html
https://workflows.tidymodels.org/reference/extract-workflow.html
https://github.com/awsm-research/Rnalytica